---
title: "R Notebook"
output:
  pdf_document: default
---

```{r install a}
list.of.packages <- c("tidyverse","ggplot2","readr","lubridate","leaflet","lattice","data.table","sqldf","RColorBrewer","scales","car","RVAideMemoire","mvShapiroTest","biotools","npmv","nnet","mlogit","pscl","boot")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages))install.packages(new.packages)

```

```{r message=FALSE}
library(readxl)
library(tidyverse)
library(ggplot2)
library(magrittr)
library(car)
library(corrplot)
library(RVAideMemoire)
library(mvShapiroTest)
library(mvnormtest)
library(npmv)
library(pscl)
library(boot)
library(biotools)
library(MASS)
```

Read in the data
```{r}
ft <- read_xlsx('chyrs.xlsx', sheet=3, col_names = TRUE)
head(ft)
```

Data Cleanup
```{r}
ft$location <- as.factor(ft$location)
ft[c("fwgt","flgt","fdiam")] <- lapply(ft[c("fwgt","flgt","fdiam")],as.double)
ft[c("seedno","lavano")] <- lapply(ft[c("seedno","lavano")],as.integer)
print(str(ft))
print(any(is.na(ft)))
```


```{r}
#remove record with na
ft <- na.omit(ft)
summary(ft)
```


Univirate analysis of the seed data

1) Fruit Length
```{r}
ggplot(data=ft,aes(x=flgt))+geom_histogram()+ ggtitle("Histogram of Fruit Length") 

ggplot(data=ft,aes(x=flgt))+geom_histogram(bins=15)+geom_freqpoly()+facet_wrap(~location,ncol=3) +ggtitle("Histogram of Fruit Length By Location") 

ggplot(data=ft,aes(x=flgt))+geom_freqpoly(bins=15)+facet_wrap(~location,ncol=3) +ggtitle("Histogram of Fruit Length By Location")

ggplot(data=ft,aes(x=location,y=flgt))+geom_boxplot()+ ggtitle("Boxplot of Fruit Length") 

ggplot(data=ft,aes(sample=flgt)) + geom_qq(distribution = stats::qnorm)+ stat_qq_line() + facet_wrap(~location)+ facet_wrap(~location)+ggtitle("QQPlot For Fruit Length By Location")

```
distributions of locations 1,3,4 and 5 appear non normal. Data from location 1 is skewed and data from location 3 has extreme values


2) Fruit Width
```{r}
ggplot(data=ft,aes(x=fwgt))+geom_histogram(bins=20)+ ggtitle("Histogram of Fruit Weight") 
ggplot(data=ft,aes(x=fwgt))+geom_histogram(bins=15)+facet_wrap(~location,ncol=3) +ggtitle("Histogram of Fruit Weight By Location") 

ggplot(data=ft,aes(x=fwgt))+geom_freqpoly(bins=15)+facet_wrap(~location,ncol=3) +ggtitle("Histogram of Fruit Weight By Location")

ggplot(data=ft,aes(x=location,y=fwgt))+geom_boxplot()+ ggtitle("Boxplot of Fruit Weight") 

ggplot(data=ft,aes(sample=fwgt)) + geom_qq(distribution = stats::qnorm)+ stat_qq_line() + facet_wrap(~location)+ facet_wrap(~location)+ggtitle("QQPlot For Fruit Weight By Location")
```
locations 1, 2 and 5 appear to deviate from normality. location 1 and 2 might be skewed


3) Fruit Diameter
```{r}
ggplot(data=ft,aes(x=fdiam))+geom_histogram(bins=20)+ ggtitle("Histogram of Fruit Diameter") 

ggplot(data=ft,aes(x=fdiam))+geom_histogram(bins=15)+facet_wrap(~location,ncol=3) +ggtitle("Histogram of Fruit Diameter By Location") 

ggplot(data=ft,aes(x=fdiam))+geom_freqpoly(bins=15)+facet_wrap(~location,ncol=3) +ggtitle("Histogram of Fruit Diameter By Location")

ggplot(data=ft,aes(x=location,y=fdiam))+geom_boxplot()+ ggtitle("Boxplot of Fruit Diameter by Location")

ggplot(data=ft,aes(sample=fdiam)) + geom_qq(distribution = stats::qnorm)+ stat_qq_line() + facet_wrap(~location)+ggtitle("QQPlot For Fruit Diameter By Location")
```
location 5 has an obvious deviation from normality. 


4) 
Number of Seeds 
Seeds variable is discrete data
```{r}
#use relative frequency since it can be a sudo for probability of the number of seeds occuring
ggplot(data=ft,aes(x=seedno))+geom_bar()+ ggtitle("Histogram of Seed No") 
ggplot(data=ft,aes(x=seedno))+geom_bar()+facet_wrap(~location,ncol=3) +ggtitle("Histogram of Seed Number By Location") 
```
Discrete data will not follow the normal distribution


Lava Summary - This is also discrete data 
```{r}
summary(ft$lavano)
#Box plot of seeds
ggplot(data=ft,aes(x=lavano))+geom_bar()+facet_wrap(~location,ncol=3) +ggtitle("Histogram of Seed Number By Location") 

```
Data has a lot of zeros because fruit should not have worms


Multivirate Analysis of fruit data
```{r}
scatterplotMatrix(ft[-c(1,5,6)],regLine = FALSE,smooth = FALSE)
title(main="Scatterplot Matrix of Fruit Attributes")

scatterplotMatrix(ft[-c(1,5,6)],groups=ft$location, regLine = FALSE,smooth = FALSE)
title(main="Scatterplot Matrix of Fruit Attributes, With Location Seperation By Color")

```
The plots with color show some attribute seperation based on color. Statistical tests will confirm if these differences are significant


Perform the following tests:
1. test if fruit characteristics are different based on location
2. if lava count is different based on location



ANOVA/MANOVA
First ensure that data do not violate ANOVA assumptions, then check for MANOVA assumptions - will not include seedno because it is discrete data 


Test For Anova Assumptions
ANOVA is robust to deviations from assumptions but the ratio between sizes must be less than 1.5

```{r}
40/33
```

1.Homogeniety of variance (homoscedacity)
List Variance of fruit characteristics
```{r}
aggregate(ft[-c(1,5,6)],by=ft[1],var,na.rm=TRUE)

```

variance for each variable is not similar, so cannot assume. Utilize Levene Test to statistically determine if variance is statistically the same or different accross locations

LeveneTest. 
```{r}
leveneTest(fwgt~location, data=ft)
leveneTest(flgt~location, data=ft)
leveneTest(fdiam~location, data=ft)
```
Fail to reject the null, variance is smilar across locations


Test Positive determinant of variance-covariance matrix
```{r}
ft[-c(1,5,6)] %>% cov() %>% det()
```

Test equality of variance covariance matrix
```{r}
x <- cbind(ft$fwgt,ft$flgt,ft$fdiam)
g <- ft$location
boxM(x,g)
```

reject null hypothesis, covariance matrices are not equal. Resolve normality and retest


fruit diameter for location 1, 2, and 5 appear to not be normal, Fruit Weight for location 1, 2, 3 appear to not be normal and fruit length for locations 1 and 3 do not appear to be normal.

Confirm with the Shapiro Wilk's test

Shapiro-Wilk Test
Univirate test
```{r}
byf.shapiro(flgt~location, data=ft) 
byf.shapiro(fwgt~location, data=ft) 
byf.shapiro(fdiam~location, data=ft) 
```


Utilize Log and BoxCox transformations, to normalize the data
Log Transformation
```{r}

ft2 <- ft[c("location","fwgt","flgt","fdiam")]
ft2["fwgt"] <- log(ft2$fwgt)
ft2["flgt"] <- log(ft2$flgt)
ft2["fdiam"] <- log(ft2$fdiam)

byf.shapiro(flgt~location, data=ft2) 
byf.shapiro(fwgt~location, data=ft2) 
byf.shapiro(fdiam~location, data=ft2) 

```

qqplot of transformed variables
```{r}
ggplot(data=ft2,aes(sample=fwgt)) + geom_qq(distribution = stats::qnorm)+ stat_qq_line() + facet_wrap(~location)+ facet_wrap(~location)+ggtitle("QQPlot For Fruit Weight By Location")

ggplot(data=ft2,aes(sample=flgt)) + geom_qq(distribution = stats::qnorm)+ stat_qq_line() + facet_wrap(~location)+ facet_wrap(~location)+ggtitle("QQPlot For Fruit Weight By Location")

ggplot(data=ft2,aes(sample=fdiam)) + geom_qq(distribution = stats::qnorm)+ stat_qq_line() + facet_wrap(~location)+ facet_wrap(~location)+ggtitle("QQPlot For Fruit Weight By Location")
```
Log transformation did not normalize the data


Utilize BoxCox transformation
Function to transform the variables using boxCox
```{r}
ft3 <- ft[c("location","fwgt","flgt","fdiam")]
powerTransform(ft3$fwgt~ft3$location)
powerTransform(ft3$flgt~ft3$location)
powerTransform(ft3$fdiam~ft3$location)
```


```{r}
ft3["fwgt"] <- bcPower(ft3$fwgt,0.2091424)
ft3["flgt"] <- bcPower(ft3$flgt,-0.8350021)
ft3["fdiam"] <- bcPower(ft3$fwgt,-0.5049158)

byf.shapiro(fwgt~location, data=ft3)
byf.shapiro(flgt~location, data=ft3)
byf.shapiro(fdiam~location, data=ft3)
```

qqplot of transformed variables
```{r}
ggplot(data=ft3,aes(sample=fwgt)) + geom_qq(distribution = stats::qnorm)+ stat_qq_line() + facet_wrap(~location)+ facet_wrap(~location)+ggtitle("QQPlot For Fruit Weight By Location")

ggplot(data=ft3,aes(sample=flgt)) + geom_qq(distribution = stats::qnorm)+ stat_qq_line() + facet_wrap(~location)+ facet_wrap(~location)+ggtitle("QQPlot For Fruit Weight By Location")

ggplot(data=ft3,aes(sample=fdiam)) + geom_qq(distribution = stats::qnorm)+ stat_qq_line() + facet_wrap(~location)+ facet_wrap(~location)+ggtitle("QQPlot For Fruit Weight By Location")
```
Box Cox also did not appear to have an impact 


Remove extreme values and test again
```{r}

ft2 <- ft
#identifiy extreme values
fwgt_out <- boxplot(ft2$fwgt~ft2$location, plot=FALSE)$out
flgt_out <- boxplot(ft2$flgt~ft2$location, plot=FALSE)$out
fdiam_out <- boxplot(ft2$fdiam~ft2$location, plot=FALSE)$out

#remove the observations
ft2 <- ft2[-which(ft2$fwgt %in% fwgt_out),]
ft2 <- ft2[-which(ft2$flgt %in% flgt_out),]
ft2 <- ft2[-which(ft2$fdiam %in% fdiam_out),]

ft3 <- ft2
```


test and qqplot of log transformed variables without extreme values
```{r}

ft2["fwgt"] <- log(ft2$fwgt)
ft2["flgt"] <- log(ft2$flgt)
ft2["fdiam"] <- log(ft2$fdiam)

#qqplot of transformed variables
byf.shapiro(fwgt~location, data=ft2)
byf.shapiro(flgt~location, data=ft2)
byf.shapiro(fdiam~location, data=ft2)

#qqplots
ggplot(data=ft2,aes(sample=fwgt)) + geom_qq(distribution = stats::qnorm)+ stat_qq_line() + facet_wrap(~location)+ facet_wrap(~location)+ggtitle("QQPlot For Fruit Weight By Location")

ggplot(data=ft2,aes(sample=flgt)) + geom_qq(distribution = stats::qnorm)+ stat_qq_line() + facet_wrap(~location)+ facet_wrap(~location)+ggtitle("QQPlot For Fruit Weight By Location")

ggplot(data=ft2,aes(sample=fdiam)) + geom_qq(distribution = stats::qnorm)+ stat_qq_line() + facet_wrap(~location)+ facet_wrap(~location)+ggtitle("QQPlot For Fruit Weight By Location")
```
Deviations from normality are very evident. Utilize Box Cox transformation without extreme values


BoxCox transformation
```{r}
powerTransform(ft3$fwgt~ft3$location)
powerTransform(ft3$flgt~ft3$location)
powerTransform(ft3$fdiam~ft3$location)
```


```{r}
#boxcox transformation
powerTransform(ft3$fwgt~ft3$location)
powerTransform(ft3$flgt~ft3$location)
powerTransform(ft3$fdiam~ft3$location)

#qqplot of transformed variables
byf.shapiro(fdiam~location, data=ft3) #1,2,5
byf.shapiro(fwgt~location, data=ft3) #1, 2, 3
byf.shapiro(flgt~location, data=ft3)

#qqplots
ggplot(data=ft3,aes(sample=fwgt)) + geom_qq(distribution = stats::qnorm)+ stat_qq_line() + facet_wrap(~location)+ facet_wrap(~location)+ggtitle("QQPlot For Fruit Weight By Location")

ggplot(data=ft3,aes(sample=flgt)) + geom_qq(distribution = stats::qnorm)+ stat_qq_line() + facet_wrap(~location)+ facet_wrap(~location)+ggtitle("QQPlot For Fruit Weight By Location")

ggplot(data=ft3,aes(sample=fdiam)) + geom_qq(distribution = stats::qnorm)+ stat_qq_line() + facet_wrap(~location)+ facet_wrap(~location)+ggtitle("QQPlot For Fruit Weight By Location")
```

Box Cox and Log transformations fail to normalize the data. Utilize One-Way Multivirate Non Parametric test and logistic regression


One-Way Multivariate Non Parametric Test

factors need to be equal,compare 2 sets of tests
1) randomly remove values to make them equal 
```{r}
l1 <- ft[sample(which(ft$location==1), 33),]
l2 <- ft[sample(which(ft$location==2), 33),]
l3 <- ft[sample(which(ft$location==3), 33),]
l4 <- ft[sample(which(ft$location==4), 33),]
l5 <- ft[sample(which(ft$location==5), 33),]
l6 <- ft[sample(which(ft$location==6), 33),]

ft4 <- rbind(l1,l2,l3,l4,l5,l6)


```

Perform the multivariate non parametric test
```{r}
nonpartest(fdiam|fwgt|flgt|seedno~location,data=ft4)
```
fruits from group 5 tend to have larger characteristics

```{r}
ssnonpartest(seedno|fdiam|fwgt|flgt~location,data=ft4)
```

fruit characteristics differ by location:

Alternative tests to carry out
1. Mixture Modelling
2. Bayesian Inference




Test for difference in lava number based on location

```{r}
model4 <- glm(lavano~location,family=poisson(link="log"),data=ft)
summary(model4)
```

```{r}
#check regression fit
#pchisq(model4$deviance, df=model4$df.residual, lower.tail=FALSE)

par(mfrow = c(2,2))
plot(model4)

```


```{r}
#Poisson Regression for Seed No
#ft["sqrtlv"] <- sqrt(ft$lavano)
model2 <- glm.nb(lavano~location,data=ft)
summary(model2)
```
lava count is not statistically different based on location


```{r}
par(mfrow = c(2,2))
plot(model2)
```






```{r}
#Poisson Regression for lavano No - data has a lot of zeros 
#model3 <- zeroinfl(lavano~location,data=ft,dist = "negbin", EM = TRUE)
#summary(model3)
```

```{r}
```

